---
title: "Wilson_Exam_3"
output: html_document
---

# 1. Load and clean FacultySalaries_1995.csv file and Re-create the graph below… 

### Setup

Load libraries and read in the csv file
``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
library(ggimage)
library(patchwork)
library(GGally)
library(ggplot2)
library(dplyr)
library(grid)
library(easystats)
library(janitor)
library(skimr)
library(caret)
library(broom)
library(purrr)
fac <- read.csv('FacultySalaries_1995.csv')
str(fac)
```

### Clean the Data
Make sure col names are clean, each col is the appropriate class, and combine chemical ID's into one column and Concentration in another.
``` {r}
fac_cleaned <- fac %>%
  clean_names() %>%
  mutate(across(starts_with("avg_"), as.numeric)) %>%
  mutate(across(where(is.character), trimws)) %>%
  drop_na()

glimpse(fac_cleaned)

fac1 <- fac_cleaned %>%
  pivot_longer(cols = c(avg_full_prof_salary, avg_assoc_prof_salary, avg_assist_prof_salary),
               names_to = 'rank',
               values_to = 'salary') %>%
  mutate(rank = case_when(rank == 'avg_full_prof_salary' ~ "Full",
                          rank == 'avg_assoc_prof_salary' ~ "Assoc",
                          rank == 'avg_assist_prof_salary' ~ "Assist",
                          T ~ rank))
glimpse(fac1)
```

### Make the Plot
``` {r}
fac1_plot <- fac1 %>%
  filter(tier %in% c('I', 'IIA', 'IIB')) %>%
  ggplot(aes(x = rank, y = salary, fill = rank)) +
  geom_boxplot() +
  theme_minimal() +
  facet_wrap(~tier) +
  labs(x = 'Rank', y = 'Salary', fill = 'Rank') +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
fac1_plot
```

# 2. Build an ANOVA model and display the summary output in your report.

``` {r}
fac_mod <- aov(data = fac1,
           formula = salary ~ state + rank + tier)

summary(fac_mod)
```

# 3. The rest of the test uses another data set. The “Juniper_Oils.csv” data. Get it loaded and take a look. Then tidy it! (show the code used for tidying in your report)

``` {r}
jun <- read.csv('Juniper_Oils.csv')
glimpse(jun)

names(x = jun)

jun_cleaned <- jun %>%
  clean_names() %>%
  pivot_longer(cols = c("alpha_pinene","para_cymene","alpha_terpineol","cedr_9_ene","alpha_cedrene","beta_cedrene","cis_thujopsene","alpha_himachalene","beta_chamigrene","cuparene","compound_1","alpha_chamigrene","widdrol","cedrol","beta_acorenol","alpha_acorenol","gamma_eudesmol","beta_eudesmol","alpha_eudesmol","cedr_8_en_13_ol","cedr_8_en_15_ol","compound_2","thujopsenal"),
               names_to = 'chemical',
               values_to = 'concentration') %>%
  mutate(concentration = as.numeric(concentration),
         years_since_burn = as.numeric(years_since_burn))

glimpse(jun_cleaned)
```

# 4. Make me a graph of the following:
 x = YearsSinceBurn
 y = Concentration
 facet = ChemicalID (use free y-axis scales)

``` {r}
jun_plot <- jun_cleaned %>%
  ggplot(aes(x = years_since_burn, y = concentration)) +
  geom_smooth() +
  facet_wrap(~chemical, scales = 'free_y') +
  theme_minimal() +
  labs(x = "Concentration", y = "Years Since Burn")
jun_plot
```

# 5. Use a generalized linear model to find which chemicals show concentrations that are significantly (significant, as in P < 0.05) affected by “Years Since Burn”. 

``` {r}
significant_chemicals <- jun_cleaned %>%
  group_by(chemical) %>%
  group_modify(~ tidy(glm(concentration ~ years_since_burn, data = .x))) %>%
  filter(term == "years_since_burn", p.value < 0.05) %>%
  select(chemical, term, estimate, std.error, statistic, p.value) %>%
  arrange(p.value)

print(significant_chemicals)
```

# 6. Commit and push all your code and files to GitHub. I’ll pull your repository and grade what I find in your html report. ####

### DONE!